### 2018-06-05 latex/egsr17-cgf/main-body.tex
minor revision

\title[Draft]%
{
%Evaluation of Higher Order Diffusion against Variable Eddington Factor Methods for Rendering Participating Media
$P_N$-Method for Multiple Scattering in Participating Media
}

@@

\author
{
%Compiled on \today
paper1027
%David Koerner$^{1,2,3}$
%\qquad Jan Nov\'ak$^{1}$
%\qquad Peter Kutz$^{2}$
%\qquad Ralf Habel$^{2}$
%\qquad Wojciech Jarosz$^{1,4}$
\\
% For Computer Graphics Forum: Please use the abbreviation of your first name.
%$^1$Disney Research
%\qquad $^2$Walt Disney Animation Studios
%\qquad $^3$University of Stuttgart
%\qquad $^4$Dartmouth College
}

@@

\begin{abstract}
%Classical diffusion approximation (CDA) is very competitive for rendering highly scattering participating media, because in these scenarios, Monte Carlo based methods suffer from high computational cost. Variable Eddington Factor methods, such as flux-limited diffusion (FLD), have shown to further improve the results from CDA. In this paper, we look at CDA as a degenerate case of the $P_1$-method and investigate the open question, if going higher order gives a viable alternative to FLD. We derive the time-independent real-valued $P_N$-equations and introduce a framework for generating finite difference stencils directly from the computer algebra representation of the $P_N$-equations for arbitrary order. We further present a staggered-grid $P_N$-solver and compare it against FLD for various standard problems.
Rendering highly scattering participating media using brute force path tracing is a challenge. The diffusion approximation reduces the problem to solving a simple linear partial differential equation. Flux-limited diffusion introduces nonlinearities to alleviate the approximation error but introduces several ad-hoc assumptions. Both methods are based on the spherical harmonics expansion of the radiance field, that is truncated after the first order. In this paper, we investigate the open question of whether higher orders provide a viable alternative to these two approaches. Increasing the order introduces a set of increasingly complex coupled partial differential equations, whose growing number and complexity make them very difficult to work with. We introduce a computer algebra framework for representing and manipulating the underlying mathematical equations and use it to derive the time-independent real-valued $P_N$-equations for arbitrary orders. We further present a staggered-grid $P_N$-solver and generate its stencil code directly from the expression tree of the $P_N$-equations. Finally, we discuss how our method compares against prior work for various standard problems. We will release our computer algebra system, solver, and data as open source to ensure reproducibility of all of our results.
%\begin{classification} % according to http://www.acm.org/class/1998/
%\CCScat{Computer Graphics}{I.3.7}{Three-Dimensional Graphics and Realism}{Raytracing}
%\end{classification}
\end{abstract}

@@

>>>>

\title%
{
%Evaluation of Higher Order Diffusion against Variable Eddington Factor Methods for Rendering Participating Media
$P_N$-Method for Multiple Scattering in Participating Media
}

@@

\author[D. Koerner \& J. Portsmouth \& W. Jakob]
{
%Compiled on \today
%paper1027
David Koerner$^{1}$
\qquad Jamie Portsmouth$^{2}$
\qquad Wenzel Jakob$^{3}$
%\qquad Ralf Habel$^{2}$
%\qquad Wojciech Jarosz$^{1,4}$
\\
% For Computer Graphics Forum: Please use the abbreviation of your first name.
$^1$University of Stuttgart
\qquad $^2$Solid Angle
\qquad $^3$\'{E}cole polytechnique f\'{e}d\'{e}rale de Lausanne (EPFL)
%\qquad $^4$Dartmouth College
}

@@

\begin{abstract}
%Classical diffusion approximation (CDA) is very competitive for rendering highly scattering participating media, because in these scenarios, Monte Carlo based methods suffer from high computational cost. Variable Eddington Factor methods, such as flux-limited diffusion (FLD), have shown to further improve the results from CDA. In this paper, we look at CDA as a degenerate case of the $P_1$-method and investigate the open question, if going higher order gives a viable alternative to FLD. We derive the time-independent real-valued $P_N$-equations and introduce a framework for generating finite difference stencils directly from the computer algebra representation of the $P_N$-equations for arbitrary order. We further present a staggered-grid $P_N$-solver and compare it against FLD for various standard problems.
Rendering highly scattering participating media using brute force path tracing is a challenge. The diffusion approximation reduces the problem to solving a simple linear partial differential equation. Flux-limited diffusion introduces nonlinearities to alleviate the approximation error but introduces several ad-hoc assumptions. Both methods are based on the spherical harmonics expansion of the radiance field, that is truncated after the first order. In this paper, we investigate the open question of whether higher orders provide a viable alternative to these two approaches. Increasing the order introduces a set of increasingly complex coupled partial differential equations, whose growing number and complexity make them very difficult to work with. We \rev{use} a computer algebra framework for representing and manipulating the underlying mathematical equations and use it to derive the time-independent real-valued $P_N$-equations for arbitrary orders. We further present a staggered-grid $P_N$-solver and generate its stencil code directly from the expression tree of the $P_N$-equations. Finally, we discuss how our method compares against prior work for various standard problems. We will release our computer algebra system, solver, and data as open source to ensure reproducibility of all of our results.
%\begin{classification} % according to http://www.acm.org/class/1998/
%\CCScat{Computer Graphics}{I.3.7}{Three-Dimensional Graphics and Realism}{Raytracing}
%\end{classification}
\end{abstract}

@@

% --------------------------------------------------------------------------------------------------
\vspace{-0.1in}
\section*{Acknowledgements}
\label{sec:acknowledgements}
\vspace{-0.05in}
We thank the anonymous reviewers for their valuable and encouraging comments and feedback. We also thank Martin Frank and Benjamin Seibold for the very valuable discussions and answers to our questions. We thank Bernd Eberhardt for feedback and support. This project has been partially funded by the MSC-BW project.

### 2018-06-05 latex/egsr17-cgf/section_car.tex
minor revision

Our computer algebra representation is based on mathematical expression trees, which are hierarchies of expression instances. Those can be basic primitives, such as numbers, variables or special symbols, such as the Kronecker delta or imaginary number. Other expression types, such as integrals, derivatives, sums, products and functions build the hierarchy by referencing child expressions.

Our framework further provides manipulators, which can be executed on expressions. We implemented all manipulations, which were required for the derivation of the $P_N$-equations. This includes operations, such as application of the distributive law, substitution, constant folding, reordering of nested integrals, application of identities up to more complex operations, such as factorization of unknowns and discretization of spatial variables. The derivation steps of the real-valued $P_N$-equations in the supplemental material were all carried out using our framework.

Finally, frontends allow rendering the expression tree into different forms. We implemented frontends for rendering expression trees to \LaTeX~and C++ source code. The equations in the supplemental material were almost all rendered by the former. The latter was used for generating the stencil code used by our solver.

Note that our framework is different from a computer algebra system (CAS), in that our framework only deals with representation and manipulation of expressions, and is not concerned with finding answers to mathematical problems.

>>>>

\rev{We use a computer algebra representation for working with the $P_N$-equations. It represents the equations using a tree of mathematical expressions, which represent numbers, symbols and other expression types, such as integrals, derivatives, sums, products and functions. Further, manipulators can be executed on these expression trees to perform substitution, constant folding, reordering of nested integrals, application of identities and more complex operations. Finally, frontends allow rendering the expression tree into different forms, such as \LaTeX~and C++ source code. While we ended up implementing our own lightweight framework, off-the-shelf packages, such as SymPy (www.sympy.org), exist and would be equally suitable for our use case.}

\begin{figure}[h]
%\vspace{1in}
\centering
\includegraphics[width=0.9\columnwidth]{figures/fig_car.pdf}
\vspace{-0.16in}
\icaption{\rev{A} computer algebra framework allows to represent equations as mathematical expressions trees. It further provides a set of functions for manipulating the tree according to valid mathematical operations, such as the binomial expansion above. Frontends allow generation of source code from the expression tree.}
\label{fig:staggeredgrid}
\end{figure}

\rev{Using the computer algebra representation, we perform the derivation steps required to arrive at the real-valued $P_N$-equations (the derivation steps shown in the supplemental material were almost all rendered from the expression tree). More importantly, we use the representation to perform automated discretization and generate the stencil code used by our solver. This is detailed in section~\ref{sec:solver_precomputation}}.

### 2018-06-05 latex/egsr17-cgf/solver.tex
pre_finalization
minor revision

@@

@@

@@

\begin{wrapfigure}{r}{0.6\columnwidth}
%\begin{center}
\hspace{-.1in}
\includegraphics[width=0.6\columnwidth]{figures/fig_staggered_interpolation.pdf}
\vspace{-.1in}
%\end{center}
\end{wrapfigure}Our solver supports placement of coefficients at arbitrary staggered grid locations (see Figure~\ref{fig:staggeredgrid}). This means, that during the discretization step, the discrete location $(i, j, k)$ (at which the unknown is meant to be evaluated) might not coincide with the location of the unknown. To solve this, depending on how the two are located to each other, the parser returns an expression, which interpolates the coefficients at $(i, j, k)$ from their defined locations. It those happen to coincide, the coefficient itsself is returned. This is also done for RTE parameters, such as $\sigma_t$ or $p^{l,m}$, which are always located at the voxel center.

@@

The number of rows is determined by the number of voxels times the number of coefficients per voxel (see Figure~\ref{fig:matrix_layout}) and can therefore become very large for high resolution and high truncation order. The matrix $A$ is square and extremely sparse, due to the local structure of the finite differences discretization. Unfortunately, it is non-symmetric due to the transport term and not diagonal dominant, which rules out many standard methods for solving linear systems. We adress this by solving the normal form $A^TA\vec{u}=A^T\vec{Q}$ instead. This gives a symmetric and positive definit system matrix $A^TA$, albeit with a higher condition number. Investigation of other solution schemes (e.g. multigrid) would be an interesting avenue for future work.

>>>>

\vspace{-0.75in}

@@

\vspace{0.6in}

@@

\vspace{0.2in}

@@

\begin{wrapfigure}{r}{0.6\columnwidth}
%\begin{center}
\hspace{-.1in}
\includegraphics[width=0.6\columnwidth]{figures/fig_staggered_interpolation.pdf}
\vspace{-.1in}
%\end{center}
\end{wrapfigure}Our solver supports placement of coefficients at arbitrary staggered grid locations (see Figure~\ref{fig:staggeredgrid}). This means, that during the discretization step, the discrete location $(i, j, k)$ (at which the unknown is meant to be evaluated) might not coincide with the location of the unknown. To solve this, depending on how the two are located to each other, the parser returns an expression, which interpolates the coefficients at $(i, j, k)$ from their defined locations. If those happen to coincide, the coefficient itsself is returned. This is also done for RTE parameters, such as $\sigma_t$ or $p^{l,m}$, which are always located at the voxel center.

@@

The number of rows is determined by the number of voxels times the number of coefficients per voxel (see Figure~\ref{fig:matrix_layout}) and can therefore become very large for high resolution and high truncation order. The matrix $A$ is square and extremely sparse, due to the local structure of the finite differences discretization. Unfortunately, it is non-symmetric due to the transport term and not diagonal dominant, which rules out many \rev{iterative} methods for solving linear systems. \rev{Iterative methods are useful, as they allow balancing accuracy against performance by tweaking the convergence threshold.} We adress this by solving the normal form $A^TA\vec{u}=A^T\vec{Q}$ instead. This gives a symmetric and positive definit system matrix $A^TA$, albeit with a higher condition number. Investigation of other solution schemes (e.g. multigrid) would be an interesting avenue for future work. \rev{However, more importantly, in the presence of vacuum regions, the matrix $A$ becomes singular and the system cannot be solved at all. This requires the introduction of a minimum threshold for the extinction coefficient $\sigma_t$.}
